Scan-Centric, Frequency-Based Method for Characterizing Peaks from Direct Injection Fourier transform Mass Spectrometry Experiments

### 2022-05-18 doc/peakcharacterization_manuscript.Rmd
a few more edits in

::: {custom-style="MDPI_1.8_keywords"}
**Keywords**: Fourier-transform mass spectrometry
:::

@@

::: {custom-style="MDPI_3.1_text"}
To motivate our solution, we generated peak lists using the peak exporting functionality in Xcalibur as well as our scan-centric peak characterization, and found matching Xcalibur peaks to our assignments.
As an example, in `r figure_count$label_text("motivation")`, we show the Xcalibur calculated intensities from the simplistically averaged spectrum and theoretical peak intensities based on NAP for four peaks assigned to threonine.
As the peak intensity decreases, the deviation between the Xcalibur calculated intensity and its corresponding NAP-based theoretical intensity become larger.
This becomes more apparent when plotting the two sets of intensities directly against each other, as shown in `r figure_count$label_text("motivation")`B.
:::

@@

::: {custom-style="MDPI_5.2_figure"}
```{r centroided_peaks, dn_id = figure_count}
# figure showing fit of quadratic model to actual data points
tar_load(peak_fit_plot)
peak_fit_plot
```
:::

@@

::: {custom-style="MDPI_5.2_figure"}
```{r breaking_regions, dn_id = figure_count}
# figure showing multiple "peaks" in what is initially a single region
tar_load(split_region_plot)
split_region_plot
```
:::

@@

::: {custom-style="MDPI_5.2_figure"}
```{r intensity_scan, dn_id = figure_count}
# figure showing peak intensity correlation with m/z, intensity varying with m/z
tar_load(intensity_scan_cor_plot)
intensity_scan_cor_plot
```
:::

@@

::: {custom-style="MDPI_3.1_text"}
We have previously described the presence of high peak density (HPD) artifacts in FT-MS spectra [@mitchellNewMethodsIdentify2018].
Ideally, the peak characterization procedure should reduce and mitigate their presence and effect in the resultant reported peaks.
Their presence should be minimized by removing noise peaks, and removing peak regions that have multiple reported peaks from the same scan.
However, we expect they may also present as characterized peaks that have larger than expected frequency level standard deviations (FSD) when calculated across scans.
These peaks can be detected by simply examining the distribution of FSDs and marking outlier peaks.
To verify the mitigation of HPD regions, we converted centroided m/z's from Xcalibur to frequency values using the previously calculated model for that sample, and measured peak density to detect HPD regions, and compare them with the scan-centric peaks and FSD outliers.
Xcalibur peak density was measured using a sliding window ten points wide and a stride of one point.

@@

To convert m/z back into frequency, we can use a similar model without the roots, as well as an extra simple linear term that doesn't have an equivalent in the above frequency model.
:::

>>>>

::: {custom-style="MDPI_1.8_keywords"}
**Keywords**: Fourier-transform mass spectrometry, orbitrap, scan-centric peak characterization, Fellgett's advantage, interferograms, frequency spectrum
:::

@@

::: {custom-style="MDPI_3.1_text"}
To motivate our solution, we generated peak lists using the peak exporting functionality in Xcalibur as well as our scan-centric peak characterization, and found matching Xcalibur peaks to our assignments.
As an example, in `r figure_count$label_text("motivation")`, we show the Xcalibur calculated intensities from the simplistically averaged spectrum and theoretical peak intensities based on NAP for four peaks matching threonine.
As the peak intensity decreases, the deviation between the Xcalibur calculated intensity and its corresponding NAP-based theoretical intensity becomes larger.
This becomes more apparent when plotting the two sets of intensities directly against each other, as shown in `r figure_count$label_text("motivation")`B.
:::

@@

::: {custom-style="MDPI_5.2_figure"}
```{r centroided_peaks, dn_id = figure_count}
# showing fit of quadratic model to actual data points
tar_load(peak_fit_plot)
peak_fit_plot
```
:::

@@

::: {custom-style="MDPI_5.2_figure"}
```{r breaking_regions, dn_id = figure_count}
# showing multiple "peaks" in what is initially a single region
tar_load(split_region_plot)
split_region_plot
```
:::

@@

::: {custom-style="MDPI_5.2_figure"}
```{r intensity_scan, dn_id = figure_count}
# showing peak intensity correlation with m/z, intensity varying with m/z
tar_load(intensity_scan_cor_plot)
intensity_scan_cor_plot
```
:::

@@

::: {custom-style="MDPI_3.1_text"}
We have previously described the presence of high peak density (HPD) artifacts in FT-MS spectra [@mitchellNewMethodsIdentify2018].
Ideally, the peak characterization procedure should reduce and mitigate their presence and effect in the resultant reported peaks.
Their presence should be minimized by removing noise peaks, and removing peak regions that have multiple reported peaks from the same scan.
However, we expect they may also present as characterized peaks that have larger than expected frequency level standard deviations (FSD) when calculated across scans.
These peaks can be detected by simply examining the distribution of FSDs and marking outlier peaks.
To verify the mitigation of HPD regions, we converted centroided m/z's from Xcalibur to frequency values using the previously calculated regression model for that sample, measured peak density to detect HPD regions, and compared them with the scan-centric peaks and FSD outliers.
Xcalibur peak density was measured using a sliding window ten points wide and a stride of one point.

@@

To convert m/z back into frequency, we can use a similar model without the roots, as well as an extra simple linear term that does not have an equivalent in the above frequency model.
:::
